---
title: "To what extent is the growth of nestling blue tits (_Cyanistes caeruleus_) influenced by competition with siblings?"
author:
  - name: Richard J. Telford
    email: richard.telford@uib.no
    affiliation: University of Bergen
    footnote: Corresponding Author
address:
  - code: University of Bergen
    address: Department of Biological Sciences, University of Bergen and Bjerknes Centre for Climate Research, Post Box 7803, N-5020 Bergen, Norway
date: "`r format(Sys.Date(), '%d %B %Y')`"
header-includes:
   - \usepackage{lineno}
   - \linenumbers
output:
  bookdown::pdf_book:
    number_sections: true
documentclass: article
classoption: a4paper
bibliography: extra/blue_tits2.bib
csl: extra/elsevier-harvard_rjt.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = TRUE, warning = TRUE, message = TRUE, error = TRUE)

knitr::knit_hooks$set(inline = function(x) {
  knitr:::format_sci(x, 'md')
})

#set ggplot theme
theme_set(theme_bw(base_size = 12))
options(digits = 2)

# loadd
loadd(treat1_fx)
```

# Methods

Data were checked for inconsistencies. One nestling was reassigned to the correct year. 
Data integrity were checked by examining the distribution of the first (typically only) digit after the decimal point for day 14 tarsus length and mass for each measurer.
Differences from a uniform distribution were tested for with a chi-squared test.


The relationship between fledgling mass and net rearing manipulation was tested with mixed effect models using the `lme4` package v. `r packageVersion("lme4")` [@lme4] in R v. `r getRversion()` [@R].
The unmanipulated reference broods were used for model selection to determine which co-variables should be included in the model in addition to brood size.
Variables tested for inclusion were chick sex, hatch year and rear area. Hatch year was treated as a factor.
In the reference brood models, a unique brood identifier was used as a random effect and AIC was used to select the best performing candidate model. 
For the manipulated broods, the co-variables selected by the reference model were used together with unique identifiers for both clutch and brood as random effects. 
This final model was fitted with restricted maximum likelihood, others were fitted with maximum likelihood to allow comparison of the fixed effects. 

R packages drake v. `r packageVersion("drake")` [@drake], 
bookdown `r packageVersion("bookdown")` [@bookdown1] and
renv v. `r packageVersion("renv")` [@renv] were used to ensure reproducibility.

Code to replicate all the analyses is archived at [https://github.com/richardjtelford/many_eco_evo](https://github.com/richardjtelford/many_eco_evo).

# Results

## Last digits

Measurer 2 had an unexpected distribution of digits after the decimal point for mass (Fig. \@ref(fig:last-digit)), with an excess of zeros and deficit of threes. 
A chi-squared test find that this distribution is significantly different from a uniform distribution (p = `r readd(last_digit_chi) %>% filter(day14_measurer == 2, variable == "Mass") %>% pull(p.value) %>% signif(1)`).

```{r last-digit, fig.cap = "Distribution of first digit after the decimal point for mass and tarsus length on day 14 by measurer"}
readd(last_digit_plot)
```

This pattern is not repeated in the tarsus length data. Its cause is unclear.

## Reference broods

Models for the reference data with brood size together with co-variables sex and year had the lowest AIC \@ref(tab:ref-aic).

```{r ref-aic}
readd(ref_AIC) %>% 
  knitr::kable(caption = "Fixed effects and AIC of mixed effect models fitted to reference data.")
```


## Manipulated broods

The effect of an additional nestling on mass at 14 days is `r treat1_fx %>% filter(term == "net_rearing_manipulation") %>% pull(estimate)` $\pm$ `r treat1_fx %>% filter(term == "net_rearing_manipulation") %>% pull(std.error)` g. 
Table \@ref(tab:fixed-tab) shows effect of the co-variables.

```{r fixed-tab}
treat1_fx %>%
  filter(effect == "fixed") %>% 
  select(-group, -effect) %>% 
  knitr::kable(caption = "Fixed effects estimates and standard errors")
```

Analysis of the residuals of this model showed no gross outliers (Fig. \@ref(fig:resid)).

```{r resid, fig.cap="Residuals against fitted values for the final model."}
plot(readd(treat1))
```



# References
